---
title: "ASSR/VRA Demo"
author: "Niklas Edvall (niklas.edvall@ki.se)"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

library(lme4)
library(tidyr)
library(ggplot2)
library(psych)
library(tidyverse)
library(writexl)
library(readxl)
```

## ASSR & VRA-analys
Suggested analysis of ASSR & VRA data investigating predictive power of ASSR for future VRA thresholds in children.

## Example data (excluded from knit)

n = 40 (ears)

Simulated data according to hypothesis of linear relationship between ASSR and VRA thresholds. VRA measured at two timepoints.

```{r, include = FALSE}
ASSR <- setNames(data.frame(matrix(ncol = 7, nrow = 0)), c("ID", "ASSR_500", "ASSR_1000", "ASSR_2000", "ASSR_4000", "Age", "Sex"))
VRA <- setNames(data.frame(matrix(ncol = 5, nrow = 0)), c("ID", "VRA_500", "VRA_1000", "VRA_2000", "VRA_4000"))

VRA2 <- setNames(data.frame(matrix(ncol = 5, nrow = 0)), c("ID", "VRA2_500", "VRA2_1000", "VRA2_2000", "VRA2_4000"))

n_ears = 40

#Loop through simulating data points
for (i in c(1:n_ears))
  
{
#randomize 4 threshold values
a <- round(rnorm(n = 1, mean = 40, sd = 10), 1)
b <- round(rnorm(n = 1, mean = 40, sd = 10), 1)
c <- round(rnorm(n = 1, mean = 40, sd = 10), 1)
d <- round(rnorm(n = 1, mean = 40, sd = 10), 1)

r <- round(rnorm(n = 1, mean = 10, sd = 15), 1) #random linear shift
  
ASSR[i,1] <- i #incremental ID

ASSR[i,2] <- a
ASSR[i,3] <- b
ASSR[i,4] <- c
ASSR[i,5] <- d

#Age and sex for ASSR
age <- round(rnorm(n = 1, mean = 10, sd = 3), 0)
if (age < 0) {age <- (age * -1)} #Age positive only

ASSR[i, 6] <- age
ASSR[i, 7] <- sample(c("M", "F"), 1)

VRA[i,1] <- i #incremental ID

VRA[i,2] <- a + r
VRA[i,3] <- b + r
VRA[i,4] <- c + r
VRA[i,5] <- d + r

r <- round(rnorm(n = 1, mean = 10, sd = 15), 1) #new random linear shift for second VRA measurement

VRA2[i,1] <- i #incremental ID

VRA2[i,2] <- a + r
VRA2[i,3] <- b + r
VRA2[i,4] <- c + r
VRA2[i,5] <- d + r

rm(a,b,c,d,r,age)

}

#Pivot to long table format for plots
ASSR_long <- pivot_longer(ASSR, cols = starts_with("ASSR"))
VRA_long <- pivot_longer(VRA, cols = starts_with("VRA"))

VRA2_long <- pivot_longer(VRA2, cols = starts_with("VRA2"))

#Set factor order
ASSR_long$name <- factor(ASSR_long$name,levels = c("ASSR_500", "ASSR_1000", "ASSR_2000", "ASSR_4000"))
VRA_long$name <- factor(VRA_long$name,levels = c("VRA_500", "VRA_1000", "VRA_2000", "VRA_4000"))

VRA2_long$name <- factor(VRA2_long$name,levels = c("VRA2_500", "VRA2_1000", "VRA2_2000", "VRA2_4000"))

#Save VRA and ASSR dataframes as excel for demo

write_xlsx(ASSR, path = "ASSR_data.xlsx")
write_xlsx(VRA, path = "VRA_data.xlsx")
write_xlsx(VRA2, path = "VRA2_data.xlsx")

write.csv(ASSR, file = "ASSR_data.csv", row.names = F)
write.csv(VRA, file = "VRA_data.csv", row.names = F)
write.csv(VRA2, file = "VRA2_data.csv", row.names = F)

```

## Load data from xlsx or csv

```{r}

VRA <- read_excel("VRA_data.xlsx")
VRA2 <- read_excel("VRA2_data.xlsx")
ASSR<- read_excel("ASSR_data.xlsx")

VRA <- read.csv("VRA_data.csv")
VRA2 <- read.csv("VRA2_data.csv")
ASSR <- read.csv("ASSR_data.csv")

```

## Overview of example data
```{r}
head(ASSR)

head(VRA)

table(ASSR$Sex)

hist(ASSR$Age, main = "Age distribution")

```

## Boxplots using ggplot2 (Tukey)

```{r, figures-side, fig.show="hold", out.width="50%"}
#Boxplot ASSR
ggplot(data = ASSR_long, aes(x = name, y = value)) +
         geom_boxplot() +
         ylim(20, 70) +
         scale_x_discrete(labels = c('500','1000','2000', '4000')) +
         labs(title = "ASSR - Hearing thresholds", x = "Frequency (Hz)", y = "Threshold level (dBnHL)") +
         theme(plot.title = element_text(hjust = 0.5))

#Boxplot VRA
ggplot(data = VRA_long, aes(x = name, y = value)) +
         geom_boxplot() +
         ylim(20, 70) +
         scale_x_discrete(labels = c('500','1000','2000', '4000')) +
         labs(title = "VRA - Hearing thresholds", x = "Frequency (Hz)", y = "Threshold level (dBnHL)") +
         theme(plot.title = element_text(hjust = 0.5))

```

## Linear regression per PTA4 frequency

Dependent variables: VRA thresholds

Indepent variables: ASSR thresholds

```{r}
mod_ASSR_500 <- lm(VRA$VRA_500 ~ ASSR$ASSR_500)
mod_ASSR_1000 <- lm(VRA$VRA_1000 ~ ASSR$ASSR_1000)
mod_ASSR_2000 <- lm(VRA$VRA_2000 ~ ASSR$ASSR_2000)
mod_ASSR_4000 <- lm(VRA$VRA_4000 ~ ASSR$ASSR_4000)
```

## 500Hz, VRA ~ ASSR results

```{r}
summary(mod_ASSR_500)

ggplot(data = ASSR, aes(ASSR$ASSR_500, VRA$VRA_500)) +
  geom_point() +
  stat_smooth(method = lm) +

  labs(title = "500Hz threshold, linear regression", x = "ASSR Threshold (dBnHL)", y = "VRA Threshold (dBHL)") +
  theme(plot.title = element_text(hjust = 0.5))
```

## 1000Hz, VRA ~ ASSR results

```{r}
summary(mod_ASSR_1000)

ggplot(data = ASSR, aes(ASSR$ASSR_1000, VRA$VRA_1000)) +
  geom_point() +
  stat_smooth(method = lm) +

  labs(title = "1000Hz threshold, linear regression", x = "ASSR Threshold (dBnHL)", y = "VRA Threshold (dBHL)") +
  theme(plot.title = element_text(hjust = 0.5))
```

## 2000Hz, VRA ~ ASSR results

```{r}
summary(mod_ASSR_2000)

ggplot(data = ASSR, aes(ASSR$ASSR_2000, VRA$VRA_2000)) +
  geom_point() +
  stat_smooth(method = lm) +

  labs(title = "2000Hz threshold, linear regression", x = "ASSR Threshold (dBnHL)", y = "VRA Threshold (dBHL)") +
  theme(plot.title = element_text(hjust = 0.5))
```

## 4000Hz, VRA ~ ASSR results

```{r}
summary(mod_ASSR_4000)

ggplot(data = ASSR, aes(ASSR$ASSR_4000, VRA$VRA_4000)) +
  geom_point() +
  stat_smooth(method = lm) +

  labs(title = "4000Hz threshold, linear regression", x = "ASSR Threshold (dBnHL)", y = "VRA Threshold (dBHL)") +
  theme(plot.title = element_text(hjust = 0.5))
```

## Intraclass correlation for test re-test of VRA
## (ICC3,k) Two-way mixed, average measures, consistency

```{r}
#Transform VRA data by frequency for ICC analysis
VRA500 <- data.frame(VRA_500 = VRA$VRA_500, VRA2_500 = VRA2$VRA2_500)
VRA1000 <- data.frame(VRA_1000 = VRA$VRA_1000, VRA2_1000 = VRA2$VRA2_1000)
VRA2000 <- data.frame(VRA_2000 = VRA$VRA_2000, VRA2_2000 = VRA2$VRA2_2000)
VRA4000 <- data.frame(VRA_4000 = VRA$VRA_4000, VRA2_4000 = VRA2$VRA2_4000)

#calculate ICC
ICC(VRA500)
ICC(VRA1000)
ICC(VRA2000)
ICC(VRA4000)

#Specify session in VRA dataframes
VRA_long$session <- rep(1, each=nrow(VRA_long))
VRA2_long$session <- rep(2, each=nrow(VRA2_long))

#Gather VRA and VRA2 for plot
all_VRA <- rbind(
VRA_long %>% 
  mutate(name = str_replace(name, "VRA_", "")),

VRA2_long %>% 
  mutate(name = str_replace(name, "VRA2_", "")))

#Specify session as factor
all_VRA$session <- factor(all_VRA$session,levels = c("1", "2"))
 
# grouped boxplot
ggplot(all_VRA, aes(x=name, y=value, fill=session)) + 
    geom_boxplot() +
    labs(title = "VRA thresholds for session 1 & 2", x = "Frequency (Hz)", y = "VRA Threshold (dBHL)") +
    theme(plot.title = element_text(hjust = 0.5))
```

